Evolution of Methodologies in Delivering Reliable Software in the European Scientific Arena

### 2018-06-04 paper-ieee-trans.tex
Join the two paragraphs to fix review#2 comment (P1 C2 L24) about the emergence of the need of giving importance to SQA

% Comment
%%top of second column:  think about saying this more gently
Major developments were carried out throughout the years focused on providing new
functionalities, partially neglecting the adoption of software engineering methodologies, in the early days.
The released software was less tested, resulting in a high rate of
rollbacks and patches applied to production systems. Thereby, the stability and
reliability of the e--Infrastructures was often compromised, affecting
negatively their exploitation by the scientific communities.

Soon it became apparent the need for a better balance between the addition of new features
and the reliability of the released software. Consequently, a substantial part of the
software lifecycle was devoted to the improvement of the quality of the software being
delivered. At this point, evolving software engineering practices were
considered and gradually adopted to define and implement quality procedures, the organization of cross--functional teams
and deployment of pilot testbeds.

% Comment
%P1 C2 L18: Claiming that stability and reliability was regularly compromised is quite a bold statement -- it needs at least some citations, and preferably some explanation of what forms this compromise took for a few prominent cases.
Analysing the EC--funded software projects over the last decade,
one can observe a continuous increase in the prominence and robustness of software
testing and validation procedures. The
% Comments
%P1 C2 L24: Was this realisation imposed "from above" e.g. by coordinated EC/ERC oversight, or was it "emergent" from the demands of the independent scientific communities?
evolution of the software engineering methodologies (creation of the Agile
Manifesto \cite{agile-manifesto}, and rise of DevOps practices \cite{zhu}) together
with the parallel advancements in the information and communication technology
field (such as virtualization, the enabling technology for Cloud computing and
operating--system--level virtualization known as containerization) motivated this trend. All of which
complemented by the emergence of automation and event--response tools.

>>>>

% Comment
%%top of second column:  think about saying this more gently
Major developments were carried out throughout the years focused on providing new
functionalities, partially neglecting the adoption of software engineering methodologies, in the early days.
The released software was less tested, resulting in a high rate of
rollbacks and patches applied to production systems. Thereby, the stability and
% Comment
%P1 C2 L18: Claiming that stability and reliability was regularly compromised is quite a bold statement -- it needs at least some citations, and preferably some explanation of what forms this compromise took for a few prominent cases.
reliability of the e--Infrastructures was often compromised, affecting
negatively their exploitation by the scientific communities.
% Comments
%P1 C2 L24: Was this realisation imposed "from above" e.g. by coordinated EC/ERC oversight, or was it "emergent" from the demands of the independent scientific communities?
It became then apparent the need for a better balance between the addition of new features
and the reliability of the released software. Consequently, incoming projects progressively
devoted a substantial part of the software lifecycle to the improvement of the quality of the
software being delivered. At this point, evolving software engineering practices were considered
and gradually adopted to define and implement quality procedures, the organization of
cross--functional teams and deployment of pilot testbeds.

Analysing the EC--funded software projects over the last decade,
one can observe a continuous increase in the prominence and robustness of software
testing and validation procedures. The
evolution of the software engineering methodologies (creation of the Agile
Manifesto \cite{agile-manifesto}, and rise of DevOps practices \cite{zhu}) together
with the parallel advancements in the information and communication technology
field (such as virtualization, the enabling technology for Cloud computing and
operating--system--level virtualization known as containerization) motivated this trend. All of which
complemented by the emergence of automation and event--response tools.
